Pretty darn good control: when are approximate solutions better than approximate models

### 2023-01-26 manuscript/manuscript.Rmd
began figure action-X and the discussion leading up to it

In Fig. \ref{fig:timeseries} we show one example time series for each of the three controls we analyzed: 
the left panel uses an optimalconstant mortality control $F^*X$, the middle panel uses $0.8 F^* X$, while the right panel is controlled by the trained DRL agent.
We see that that optimal constant mortality control leads to a near-extinction event at around $t=150$.
The constant mortality policy at $0.8 F^*$ is already conservative enough to support a sustainable path.
There, the population levels are maintained roughly in a stationary state.
Finally, the DRL time series shows an oscillatory behavior, with a slight dephase between the $Y$ and $Z$ curves, while the $X$ oscillations are roughly anticorrelated with $Z$.

@@

This time series is representative of the patterns observed in the data (see App. \ref{app:tbd}).
The episode reward and episode length statistics are shown in Fig. \ref{fig:rew_len_fig}.
There, we highlight two things:
First, that the tendency towards early episode ends is fixed by the conservative approach of fishing at 80% of the optimal mortality rate.
In App. \ref{app:tbd} we discuss a warning about this approach, however: the arbitrarily chosen percentange itself lives on a ‘‘knife's edge.’’
More concretely, choosing a higher percentage, say 90%, can switch the behaviour of the controlled system and give rise to the proliferation of early episode ends.
Second, that while the conservative 80% approach has a stable behavior, it consistently produces smaller harvests than a DRL generated policy.

@@

```{r rew_len_fig, fig.cap="TBD.", out.width = "6.5in"}
library(patchwork)

>>>>

We evaluated each of the three control approaches (MSY constant mortality, 80% MSY constant mortality and DRL) by running 50 episodes using the respective control rule.
In Fig. \ref{fig:timeseries} we show the time series of one such repetition.
The left panel uses an MSY constant mortality control $F^*X$, the middle panel uses $0.8 F^* X$, while the right panel is controlled by the trained DRL agent.
We see that that optimal constant mortality control leads to a near-extinction event at around $t=150$.
The constant mortality policy at $0.8 F^*$ is already conservative enough to support a sustainable path.
There, the population levels are maintained roughly in a stationary state.
Finally, the DRL time series shows an oscillatory behavior, with a slight dephase between the $Y$ and $Z$ curves, while the $X$ oscillations are roughly anticorrelated with $Z$.

@@

This time series is representative of the patterns observed in the data (see App. \ref{app:tbd}).
The episode reward and episode length statistics are shown in Fig. \ref{fig:rew_len_fig}.
There, we highlight two things:
First, that the tendency towards early episode ends is fixed by the conservative approach of fishing at 80% of the optimal mortality rate.
(In App. \ref{app:tbd} we discuss a warning about this approach, however: the arbitrarily chosen percentage itself lives on a ‘‘knife's edge.’’
More concretely, choosing a higher percentage, say 90%, can switch the behaviour of the controlled system and give rise to the proliferation of early episode ends.)
Second, that while the conservative 80% approach has a stable behavior, it consistently produces smaller harvests than a DRL generated policy.

```{r} 
#| label: rew_len_fig
#| fig.cap: "Bar charts for the episode lengths (top panel) and episode total harvests (bottom panel) for 50 episodes." 
#| out.width: "6.5in"
p1 <- msy_tac_ppo |> 
    group_by(rep, policy) |>
    filter(t==max(t), species == "X", rep < 50) |>
    ggplot(aes(rep, t)) + 
      geom_col() +
      facet_wrap(~factor(policy, levels = c("msy", "tac", "ppo")), labeller = policy_labeller ) +
      labs(y = "episode length", x = "repetition") +
      theme(axis.text.x = element_blank())

To understand the DRL-derived policy, Fig. \ref{fig:action-X} depicts the harvest as a function of $X$.
Specifically, for each point $(x,y,z)$ visited by one of the aforementioned 50 evaluation episodes, we evaluate the harvest collected by the DRL agent and we subsequently plot the scatter data as a function of the projection $x$ onto the $X$ axis.

```{r}
#| label: action-X
#| fig.cap: "TBD"
#| out.width: "6.5in"
p <- ppo_df |>
  mutate(X = X+1, Y = Y+1, Z = Z+1, harvest = action*X)

```

@@

```{r rew_len_fig_old, fig.cap="TBD.", out.width = "6.5in"}
library(patchwork)
